---
title: "ExeRcices Metsovo"
author: "Timothée Giraud"
date: "25 janvier 2016"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: journal
    toc: yes
---


# Install and load a package
Install and load the package named `cartography`.
=> search for how to do it on internet


```{r}
# install.packages("cartography")
library(cartography)
```


# Load some data

```{r}
data(nuts2006)
```

nuts2006 is a dataset embedded in the cartography package.
It contains different objects:
- data frame (ending with df)  
- spatial data frames sobjects (ending with .spdf)  

You can click on a dataframe to view it on RStudio.
You can also show the first lines of the data frame on the console with the `head` function.
```{r}
head(nuts2.df)
```

Try to show only the 6 firsts lines on the console.
type ?head on the console to show function help in RStudio or search for "R head" on the internet.

You can plot (display) a spatial object with the plot function.
```{r}
plot(nuts2.spdf)
```

You can access other attributes (data, projection...) of spatial objects with the `@`:
```{r}
# projection info 
nuts2.spdf@proj4string
# the full attribute table
nuts2.spdf@data
# first line of the attribute table
head(nuts2.spdf@data)

```







# Modify and adapt an existing code

Here is some code that plot a map of countries population in Europe: 
```{r}
library(cartography)
data(nuts2006)
plot(nuts0.spdf, col = "red",border = "blue")
# Population plot on proportional symbols
propSymbolsLayer(spdf = nuts0.spdf, df = nuts0.df,
                 var = "pop2008", k = 0.001, border = "white",
                 symbols = "circle", col =  "green",
                 legend.pos = "bottomleft", 
                 legend.title.txt = "Population in 20118", 
                 legend.style = "c")
```

The map is not very informative, circles are too small, colors are not carefully choosen.
You can adapt the map and choose better parameters to make a better map, for example you can change the colors of the basemap, make bigger symbols, modify the legend etc..  
See ?propSymbolsLayer (help of the function) to see what is possible.  
You can also check the layoutLayer function to add a layout to the map (title, scale, north arrow, author...)


# Search for the travel time between your hometown and Metsovo

The osrm package allows to get travel times and travel geometries thanks to OpenStreetMap database.


```{r}
library(osrm)

route <- osrmViaroute(src = c(39.771395, 21.181455), 
                      dst = c(39.640706, 20.854831))
route
routeGeom <- route <- osrmViarouteGeom(src = c("Metsovo",39.771395, 21.181455), 
                                       dst = c("Ioannina",39.640706, 20.854831), 
                                       sp = TRUE)
plot(routeGeom)
routeGeom@data
```

At this point it is possible to export the road to a GIS format:

```{r}
library(rgdal)
# writeOGR(obj = routeGeom, dsn = getwd(), layer = "myroad", driver = "ESRI Shapefile")
```



# Show OpenStreetMap tiles in R


```{r}
library(cartography)

library(OpenStreetMap)
osmTiles <- getTiles(spdf = routeGeom)

tilesLayer2(osmTiles)

plot(routeGeom, add=T, col = "blue", lty = 2, lwd = 2)

```





