---
title: "ExeRcises"
author: "Timothée Giraud"
date: "13 janvier 2016"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: journal
    toc: yes
---

The work done in this exercise must be presented in a webpage or in a pdf.

# Import Data in R
The first step is to import some data in R. 

## Import Statistical Data 
- Import data on countries (/data/world.csv)  
```{r}
# Use read.csv to import csv in R
world <- read.csv(file = "data/world.csv")
# Show the first lines of a data frame
head(world)
```




## Import Geographical Data
- Import the world countries base-map. Which package is needed to do so?
```{r}
# The rgdal package is the "go to" package to handle spatial data in R 
library(rgdal)
worldbasemap <- readOGR(dsn = "data/shp", layer = "countries")

# Spatial objects can be plotted
plot(worldbasemap)
worldbasemap@proj4string
```


# Get information on the dataset
- How many countries are documented in this file ?
```{r}
# Number of lines in a data frame
nrow(world)

```
- How many variables describe the countries ?
```{r}
# Number of columns
ncol(world)
```


- Which country has the greatest share of urban population ? You will have to create a variable (share of urban population) and find a way to get its maximum value (see `?order` or `?which.max` functions)

```{r}
# create a new variable
world$URBSHARE <- world$URB / world$POP * 100

# select the line with the max value
world[which.max(world$URBSHARE),]

# or order the data frame along the variable
world <- world[order(world$URBSHARE, decreasing = TRUE), ]
head(world)

```


# Manipulate the base map

- Display a map of the world (change countries colors, boundaries, background...)
```{r}
# Plots of spatial objects can be customized
plot(worldbasemap, 
     col  = "#E3DEBF", 
     border = "white", 
     lwd=0.75, 
     bg = "#A6CAE0")

```


- Change the projection of the map (see the `?spTransform` function from the `sp` package).

```{r}
# The spTransform function allow to transform projections
# => http://spatialreference.org/
robinson_proj <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
worldbasemap.reproj <- spTransform(x = worldbasemap, 
                                   CRSobj = robinson_proj)

plot(worldbasemap.reproj, 
     col  = "#E3DEBF", 
     border = "white", 
     lwd=0.75, 
     bg = "#A6CAE0")
```



# Thematic Mapping
- Create a map of the countries population with a full layout (north arrow, legend, title...) (see the  `?cartography` package)

```{r, fig.height = 3.5}
library(cartography)
opar <- par(mar = c(0,0,1.2,0))
plot(worldbasemap.reproj, 
     col  = "#E3DEBF", 
     border = "white", 
     lwd=0.75, 
     bg = "#A6CAE0")

propSymbolsLayer(spdf = worldbasemap.reproj, df = world, var = "POP", 
                 border = "grey80", col = "#920000", lwd = 0.5,
                 legend.pos = "left", legend.title.txt = 'Population', 
                 legend.values.rnd = -2)
layoutLayer(title = "World Population", 
            sources = "", 
            author = "T. Giraud, 2016", 
            scale = NULL)

par(opar)
```
- Create a map of the countries urbanization.
```{r, fig.height = 3.5}
boxplot(world$URBSHARE)
hist(world$URBSHARE)
summary(world$URBSHARE)
opar <- par(mar = c(0,0,1.2,0))

layoutLayer(title = "World Urbanization", extent = worldbasemap.reproj,
            sources = "", col = "NA",
            author = "T. Giraud, 2016", 
            scale = NULL, bg =  "#A6CAE0", 
            frame = FALSE, coltitle = "#952020")


choroLayer(spdf = worldbasemap.reproj, df = world, 
           var = "URBSHARE", 
           method = "quantile", nclass = 8, 
           col = carto.pal(pal1 = "green.pal", n1 = 4, 
                           pal2 = "wine.pal", n2 = 4), 
           border = "grey40", lwd = 0.5, legend.pos = "left", 
           legend.title.txt = "Urbanization\nrate (%)", add=TRUE) 

## An other solution could be:

plot(worldbasemap.reproj,
     col  = "#E3DEBF",
     border = "white",
     lwd=0.75,
     bg = "#A6CAE0")
propSymbolsChoroLayer(spdf = worldbasemap.reproj, df = world, var = "POP",
                      border = "grey80", lwd = 0.5, var2 = "URBSHARE",
                      method = "quantile", nclass = 8,
                      col =  carto.pal(pal1 = "green.pal", n1 = 4,
                                       pal2 = "wine.pal", n2 = 4),
                      legend.var.pos = "right", legend.var.title.txt = 'Population',
                      legend.var.values.rnd = -2, legend.var2.pos = "left",
                      legend.var2.title.txt = "Urbanization\nrate (%)" )
layoutLayer(title = "World Urbanization", 
            sources = "", col = "NA",
            author = "T. Giraud, 2016", 
            scale = NULL, 
            coltitle = "#952020")

```



# Distance calculations
- Which are the countries the more away from each other ? (see `spDists`, `which`...)
- Which are the countries the closest from each other ? 

```{r}
library(sp)
worldbasemap.WGS84 <- spTransform(x = worldbasemap.reproj, 
                                  CRSobj = "+init=epsg:4326")
mat <- spDists(x = worldbasemap.WGS84, y = worldbasemap.WGS84, 
               longlat = TRUE)
dimnames(mat) <- list(worldbasemap.WGS84$ID, worldbasemap.WGS84$ID)
diag(mat) <- NA
which(mat == max(mat, na.rm = T), arr.ind = TRUE)
which(mat == min(mat, na.rm = T), arr.ind = TRUE)

```

# Geoprocessing
- Import the layer of Souli localities and the Corine Land cover layer of Souli.


```{r}
library(rgdal)
localities <- readOGR(dsn = "data/shp", layer = "oikismoi_souli")
landcover <- readOGR(dsn = "data/shp", layer = "souli_corine")

# add the projection info 
landcover@proj4string <- localities@proj4string

head(localities@data)
head(landcover@data)
```


- Plot a map with each dot colored with the land use it lays on. (see `over` from `sp` package)

```{r}
row.names(x = localities) <- localities$OBJECTID
locCLC <- over(x = localities, y = landcover)
locCLC$ID <- row.names(locCLC)
opar <- par(mar = c(0,0,1.2,0))
plot(landcover, col = "grey80", lwd = "0.5", border = "grey50")

propSymbolsTypoLayer(spdf = localities, dfid = "ID", k = 0.001,
                     df = locCLC, 
                     var = "DXF_COLOR",legend.var.pos = "n",
                     var2 = "CODE_00")

```


- Use detailed names in the legend (see the clclegend.csv file and `?merge`)

```{r}

clcNames <- read.csv("data/clclegend.csv")

locCLC <- merge(locCLC, clcNames, by.x= "CODE_00", by.y = "CLC_CODE", all.x)
locCLC$LABEL3 <- as.character(locCLC$LABEL3)

# osm <- getTiles2(spdf = landcover, crop = T)
# tilesLayer2(osm)
plot(landcover, col = "grey80", lwd = "0.5", border = "grey50", add=F)

propSymbolsTypoLayer(spdf = localities, dfid = "ID", k = 0.001,
                     df = locCLC, 
                     var = "DXF_COLOR",legend.var.pos = "n",
                     var2 = "LABEL3")
layoutLayer()
par(opar)
```




